Supporting Information

### 2016-09-21 portfolio/text/anderson-etal-rev-pfx.tex
Start drafting results
Tweak text after Eric edits

\title{Portfolios of fishing permits and species
reduce risk but adjusting fishing strategy is risky}
\author{
Sean C. Anderson$^1$,
Eric J. Ward$^2$,
Milo D. Adkison$^3$,
Anne H. Beaudreau$^3$,\\
Richard E. Brenner$^4$,
Alan C. Haynie$^5$,
Andrew O. Shelton$^2$,\\
Jennifer C. Shriver,
Jordan T. Watson
}

>>>>

\title{Benefits of diversification and risks of 
changing strategies for individual fishers\\
Portfolios of fishing permits and species
reduce risk but adjusting fishing strategy is risky\\
Portfolio benefits to individual fishers
depend on approach to diversification\\
Portfolio benefits of income diversification
are unavailable to many fishermen\\
Effect of species diversity on income variability
for individuals depends on fishing strategy\\
Effect of species diversity on revenue risk
for individual fishers
depends on diversification approach\\
}
\author{
Sean C. Anderson$^1$,
Eric J. Ward$^2$,
Milo D. Adkison$^3$,
Anne H. Beaudreau$^3$,\\
Richard E. Brenner$^4$,
Alan C. Haynie$^5$,
Andrew O. Shelton$^2$,\\
Jennifer C. Shriver,
Jordan T. Watson,
\ldots
}

\newlabel{tab:strategies}{1}
\newlabel{fig:fe}{1}
\newlabel{fig:re}{2}
\newlabel{fig:spoke-all}{3}
\newlabel{fig:resid}{4}
\newlabel{fig:resid-downside}{5}
\newlabel{fig:str-yrs}{6}

### 2016-09-21 portfolio/text/ms.Rmd
Tweak text after Eric edits

*Eric: Title should emphasize individuals, e.g. "Benefits of diversification and risks of changing strategies for individual fishermen/fishers"*
<!-- Fishing is inherently risky and we have long-sought regulations and fishing behaviors that reduce this risk. -->
<!-- The increasing the diversity of species fished has been proposed to reduce risk for fishermen -->

@@

- there is often limited room for intentional diversification within a permit
- in fact, specializing can be profitable within some fisheries
- analog here in farming literature is that diversification
  for small scale farmers generally good, except when they
  specialize on high demand/high value crops (tobacco, cocoa). 
  Seems like benefits of diversification are also smaller for 
  large-scale industrial farmers.

@@

Natural resources such as forests, land animals, and fishes
provide a primary source of income for millions of people worldwide [REFS].
However, environmental,
biological, and price variation
combined with regulatory changes
and human-caused environmental disasters
generate high levels of income variability and risk
for these individuals.
For example, drought and flooding are major sources of risk 
to agricultural food security 
(http://rstb.royalsocietypublishing.org/content/360/1463/2139.short),

And EVOS example

@@

- crop rotation has been proposed as metric for buffering against short term
  variation and long term trends (like climate change)

BUT in some situations, external forces (prices, demand) may discourage the adoption of these diverse strategies and help promote specialization, 
http://link.springer.com/article/10.1007/s10584-004-0710-z

@@

This effect has been identified for fisheries.
At community level, ... [@sethi2014]
At vessel level, ... [@kasperski2013] ...

Most papers on fisheries have adopted the approach similar to Kasperski et
al. and used data reported to the level of vessels or higher. This analysis
is unique in that we're able to separate individual people from vessels
(enabling us to track people that switch fishery sectors over time, or
change strategies and utilize multiple vessels)

@@

*Eric: I think we need a little here on the AK data and why that's both important and generalizable to fisheries around the world. 
Fisheries in Alaska are the highest valued fisheries in N. America
with landings valued at XX $ billion annually. There are a number of fishing
strategies where both specialization and diversification are extremely profitable. There's a number of fishing permits that constrain the potential diversity of species landed (e.g. herring, halibut, sablefish, or crab permits that allow the user to only land that species) versus broader 'finfish' permits that allow users to land a wide variety of groundfish spp. 
- perhaps we also need to have a sentence or two on the implicit constraints of geography and gear / vessel size on species diversification*

@@

We obtained fisheries landings and revenue data for
all permit holders in Alaska
from `r range_yrs[1]` to `r range_yrs[2]`
from the Alaskan Commercial Fisheries Entry Commission (CFEC).
These data represent a total of
nearly `r round(ndat$initial_tot_rev/1e9)` billion 2009 USD
and a combined `r round(ndat$initial/1e6,1)` million rows of data.
We adjusted for inflation by setting all revenues in 2009 USD
based on data from TODO.
We implemented a number of filtering steps
to focus on individuals actively engaged in commercial fishing
(Supporting Information).
In particular, we removed permit holders whose median revenue
was less than $10,000 2009 USD.

We combined similar permits into permit groupings 
and identified sets of 
permit groupings held by an individual 
in a specific year as 'strategies' (Supporting Information). 
For example, we combined permits that were the same except for vessel size 
(e.g. a longliner fishing sablefish on a vessel under 60ft was 
considered to have the same strategy as a longliner fishing 
sablefish on a vessel over 60ft),
and we combined a number of permits that were targeting a single species 
and differed only in region fished.
Therefore, someone fishing herring roe in Kodiak
and someone fishing herring roe in Alaska Peninsula
were both considered to have a 'herring roe' permit group.
For salmon permits, 
we grouped permits using the same gear type across regions
for groups that shared similar species composition. 
After this process of combining permits,
we were left with `r n_grouped_permits`
unique fishing permit groups (Table S\ref{tab:strategies}),
which we aggregated
within each person-year combination 
to form permit strategies.
For example, if someone fished halibut
and sablefish in the same year,
their strategy for that year would be `halibut-sablefish'.
Specific details on our strategy definitions are available in 
the Supporting Information.

>>>>

<!-- Fishing is inherently risky and we have long-sought regulations and fishing behaviors that reduce this risk. -->
<!-- The increasing the diversity of species fished has been proposed to reduce risk for fishermen -->

@@

- there is often limited room for intentional diversification within a permit
- in fact, specializing can be profitable within some fisheries
- Eric: analog here in farming literature is that diversification
  for small scale farmers generally good, except when they
  specialize on high demand/high value crops (tobacco, cocoa). 
  Seems like benefits of diversification are also smaller for 
  large-scale industrial farmers.

@@

Natural resources such as forests, land animals, and fishes
provide a primary source of income for millions of people worldwide [REFS].
However, environmental,
biological, and price variation
combined with regulatory changes
and human-caused environmental disasters
generate high levels of income variability and risk
for these individuals.
For example, drought and flooding are major sources of risk 
to agricultural food security 
(http://rstb.royalsocietypublishing.org/content/360/1463/2139.short),
And EVOS example

@@

- crop rotation has been proposed as metric for buffering against short term
  variation and long term trends (like climate change)
- BUT in some situations, external forces (prices, demand) may discourage the
  adoption of these diverse strategies and help promote specialization, 
http://link.springer.com/article/10.1007/s10584-004-0710-z

@@

This effect has been identified for fisheries.
At community level, ... [@sethi2014]
At vessel level, ... [@kasperski2013] ...
- Eric: Most papers on fisheries have adopted the approach similar to Kasperski et
al. and used data reported to the level of vessels or higher. This analysis
is unique in that we're able to separate individual people from vessels
(enabling us to track people that switch fishery sectors over time, or
change strategies and utilize multiple vessels)

@@

*Eric: I think we need a little here on the AK data and why that's both important and generalizable to fisheries around the world. 
Fisheries in Alaska are the highest valued fisheries in N. America
with landings valued at XX $ billion annually. There are a number of fishing
strategies where both specialization and diversification are extremely profitable. There's a number of fishing permits that constrain the potential diversity of species landed (e.g. herring, halibut, sablefish, or crab permits that allow the user to only land that species) versus broader 'finfish' permits that allow users to land a wide variety of groundfish spp.* 
*perhaps we also need to have a sentence or two on the implicit constraints of geography and gear / vessel size on species diversification*

@@

We obtained fisheries landings and revenue data for
all permit holders in Alaska
from `r range_yrs[1]` to `r range_yrs[2]`
from the Alaskan Commercial Fisheries Entry Commission (CFEC).
These data represent a total of
nearly `r round(ndat$initial_tot_rev/1e9)` billion 2009 USD
and a combined `r round(ndat$initial/1e6,1)` million rows of data.
We adjusted for inflation by setting all revenues in 2009 USD
based on data from TODO.
We implemented a number of filtering steps
to focus on individuals actively engaged in commercial fishing
(Supporting Information).
In particular, we removed permit holders whose median revenue
was less than \$10,000 2009 USD.

We combined similar permits into permit groupings 
and identified sets of 
permit groupings held by an individual 
in a specific year as 'strategies' (Supporting Information). 
For example, we combined permits that were the same except for vessel size 
(e.g. a longliner fishing sablefish on a vessel under 60 ft was 
considered to have the same strategy as a longliner fishing 
sablefish on a vessel over 60 ft),
and we combined a number of permits that were targeting a single species 
and differed only in region fished.
Therefore, someone fishing herring roe in Kodiak
and someone fishing herring roe in Alaska Peninsula
were both considered to have a 'herring roe' permit group.
For salmon permits, 
we grouped permits using the same gear type across regions
for groups that shared similar species composition. 
After this process of combining permits,
we were left with `r n_grouped_permits`
unique fishing permit groups (Table S\ref{tab:strategies}),
which we aggregated
within each person-year combination 
to form permit strategies.
For example, if someone fished halibut
and sablefish in the same year,
their strategy for that year would be `halibut-sablefish'.
Specific details on our strategy definitions are available in 
the Supporting Information.

### 2016-09-21 portfolio/text/supp.Rnw
Tweak text after Eric edits

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=5in]{../figs/stan-main-effects.pdf}
\caption{Main effects from the hierarchical model
predicting revenue ($\beta$) and variability ($\gamma$) in revenue.
The $\eta$ coefficients represents strategy-level predictors
relating mean species diversity and mean days fished
within a strategy
(both centered by their means and divided by
two times their standard deivations)
to the magnitude of residual error.}
\label{fig:fe}
\end{center}
\end{figure}

>>>>

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=5in]{../figs/stan-main-effects.pdf}
\caption{Main effects from the hierarchical model
predicting revenue ($\beta$) and variability ($\gamma$) in revenue.
The $\eta$ coefficients represents strategy-level predictors
relating mean species diversity and mean days fished
within a strategy
(both centered by their means and divided by
two times their standard deivations)
to the magnitude of residual error. (TODO symbols need to be updated)}
\label{fig:fe}
\end{center}
\end{figure}
